Introduction

### 2024-11-25 finalestimate.Rmd
firstfinalpushmaybe

Comparing this model to the prior ones, we see that the reciprocity
estimate is still significantly positive (b = 1.3341, SE = 0.4543). The
parameters of the structural network statistics also show similar
patterns in estimates and significance as they did in model 1.
Concerning the added covariate "gender", we find that the parameter
estimates for both covariate-related activity (egoX) , pertaining to the
gender of the ego, and covariate-related popularity (altX), relating to
the gender of the alter, do **not** show any significance and hence do
not allow for reliable inferences. The 'same-covariate' measure (sameX)
for gender is significantly positive (b = 0.6637, SE =0.2685). It
reveals that scholars, indeed, prefer same-gender collaborators. This
supports H1, which posits gender homophily in collaboration patterns.

Model 2 provides strong evidence for gender homophily (H1), with
scholars more likely to collaborate with same-gender peers, but no
significant differences in collaboration tendencies based on the gender
of ego or alter.

@@

This model is partly similar to Model 2 but additionally holds the
network statistics egoX and altX for the main effect of department, to
investigate differences in same-gender collaboration patterns based on
the two departments.

Originally, I planned to also include an interaction term of departm_RS
and gender_RS to obtain clearer insights into variations in patterns of
gender homophily in relation to scholars' departments. However, I have
refrained from adding another model to this project. I decided against
this partly due to time constraints but I also reason that I can use the
comparison of model 2 and 3 to examine changes in the estimate of
'sameX' (gender), to *tentatively* infer whether the department of
scholars modifies the collaboration patterns at hand.

Model 3 has a maximum convergence ratio of **0.182** and a Goodness of
Fit on the indegree distribution of **0.108**.

Interestingly, the estimate for outdegree (density) that had lost its
significance in model 2, is now back to being significant and negative
(b = -4.3233, SE = 1.0871).

In terms of estimate size and significance of the structural network
statistics we, once again, find similar patterns as in models 1 and 2.
However, the outdegree parameter has now reached significance (b =
-4.3233, SE = 1.0871).

@@

When it comes to the parameters pertaining to the added covariate
'department', we find that any differences in collaboration based on the
department of an ego (egoX) or of an alter (altX) are not warranted, as
the estimates of these parameters are insignificant. However, we find a
significantly positive estimate for the parameter of 'covariate-related
identity' (sameX), which indicates that scholars are more likely to
collaborate with people from their own department, indicating
**department-based homophily**. (b = 2.7240, SE = 0.7277).

Model 3 highlights persistent gender homophily (H1) across departments
but does not provide evidence for differences in gender homophily
between Sociology and Political Science, contrary to H2.

@@

Model 4 has a maximum convergence ratio of **0.23,** and a a Goodness of
Fit on the indegree distribution of **0.197**.

**Structural network statistics.** The estimate for outdegree is
negative but not significant. This suggests that, after accounting for
other variables, the baseline likelihood of forming ties remains low and
is not a major driver of collaboration patterns. Reciprocity has a
significant positive estimate (b = 2.1770, SE = 0.5960). This reflects a
strong tendency for mutual collaborations, meaning if one scholar
collaborates with another, the latter is likely to collaborate back
(dyadic reciprocity). The paramters Indegree Popularity and Outdegree
Activity are both insignificant, suggesting that neither the tendency to
receive many collaborations (indegree popularity) nor the tendency to
initiate them (outdegree activity) is significantly relevant to the
network structure. The network isolate estimate is, once again,
significantly positive ( b = 5.4634, SE = 2.2770), indicating that as
isolates tend to remain isolated.

**Gender.** Similar to AltX and EgoX in prior models, both an alter's
gender (b = -0.0903, SE = 0.3372) and an ego's gender (b = 0.2993, SE =
0.4934) still have insignificant effects. The parameter SameX shows a
significant positive estimate (b = 0.6820, SE = 0.2804) which barely
differs from the two prior models. This suggests that adding the
covariate 'H-Index' did not prompt any major changes in scholars'
preferences for same-gendered collaborators.

**Department.** While an alter's departmen**t** (AltX) remained
insignificant to the collaboration patterns, the estimates for an ego's
department changed incomparison to the prior model. In this model, egoX
seems to be of significant relevance (b = -1.4652, SE = 0.6963), with
its negative estimate indicating that, once effects of h-index are
controlled for, scholars from the political science (coded as 1)
department have a lower tendency to initiate collaborations than
scholars from the sociology department (coded as 0).\
Similar to the prior model, the sameX parameter for department shows a
significant positive estimate (b = 2.6327, SE = 0.6080) scholars are
still more likely to collaborate with others from the same department.

**H-Index**. Both an alter's and an ego's H-Index seem to be relevant in
collaboration preferences. The significant and positive estimateof AltX
(b = 0.0829, SE = 0.0364) suggest that scholars with higher h-index are
more preferred as collaborators. The egoX estimate is significant and
negative (b = -0.1591, SE = 0.0744) suggesting that scholars with higher
h-index are less likely to initiate collaborations.

The results partially support **H4**, as the h-index is shown to
influence collaboration patterns, with prestigious scholars being more
likely to attract collaborations as alters but less likely to initiate
them as egos. However, gender homophily remains significant even after
controlling for the h-index, suggesting that prestige does not
substantially weaken the influence of gender homophily in co-authorship
networks.

>>>>

Comparing this model to the prior ones, we see that the reciprocity
estimate is still significantly positive (b = 1.3341, SE = 0.4543). The
parameters of the structural network statistics also show similar
patterns in estimates and significance as they did in model 1.
Concerning the added covariate "gender", we find that the parameter
estimates for both covariate-related activity (egoX) , pertaining to the
gender of the ego, and covariate-related popularity (altX), relating to
the gender of the alter, do **not** show any significance and hence do
not allow for reliable inferences. The 'same-covariate' measure (sameX)
for gender is significantly positive (b = 0.6637, SE =0.2685). It
reveals that scholars, indeed, prefer same-gender collaborators. This
supports H1, which posits gender homophily in collaboration patterns.
Model 2 provides strong evidence for gender homophily (H1), with
scholars more likely to collaborate with same-gender peers, but no
significant differences in collaboration tendencies based on the gender
of ego or alter.

@@

This model is partly similar to Model 2 but additionally holds the
network statistics egoX and altX for the main effect of department, to
investigate differences in same-gender collaboration patterns based on
the two departments. Originally, I planned to also include an
interaction term of departm_RS and gender_RS to obtain clearer insights
into variations in patterns of gender homophily in relation to scholars'
departments. However, I have refrained from adding another model to this
project. I decided against this partly due to time constraints but I
also reason that I can use the comparison of model 2 and 3 to examine
changes in the estimate of 'sameX' (gender), to *tentatively* infer
whether the department of scholars modifies the collaboration patterns
at hand.

Model 3 has a maximum convergence ratio of **0.182** and a Goodness of
Fit on the indegree distribution of **0.108**. Interestingly, the
estimate for outdegree (density) that had lost its significance in model
2, is now back to being significant and negative (b = -4.3233, SE =
1.0871). In terms of estimate size and significance of the structural
network statistics we, once again, find similar patterns as in models 1
and 2. However, the outdegree parameter has now reached significance (b
= -4.3233, SE = 1.0871).

@@

When it comes to the parameters pertaining to the added covariate
'department', we find that any differences in collaboration based on the
department of an ego (egoX) or of an alter (altX) are not warranted, as
the estimates of these parameters are insignificant. However, we find a
significantly positive estimate for the parameter of 'covariate-related
identity' (sameX), which indicates that scholars are more likely to
collaborate with people from their own department, indicating
**department-based homophily**. (b = 2.7240, SE = 0.7277). Model 3
highlights persistent gender homophily (H1) across departments but does
not provide evidence for differences in gender homophily between
Sociology and Political Science, contrary to H2.

@@

Model 4 has a maximum convergence ratio of **0.23,** and a a Goodness of
Fit on the indegree distribution of **0.197**. \
**Structural network statistics.** The estimate for outdegree is
negative but not significant. This suggests that, after accounting for
other variables, the baseline likelihood of forming ties remains low and
is not a major driver of collaboration patterns. Reciprocity has a
significant positive estimate (b = 2.1770, SE = 0.5960). This reflects a
strong tendency for mutual collaborations, meaning if one scholar
collaborates with another, the latter is likely to collaborate back
(dyadic reciprocity). The paramters Indegree Popularity and Outdegree
Activity are both insignificant, suggesting that neither the tendency to
receive many collaborations (indegree popularity) nor the tendency to
initiate them (outdegree activity) is significantly relevant to the
network structure. The network isolate estimate is, once again,
significantly positive ( b = 5.4634, SE = 2.2770), indicating that as
isolates tend to remain isolated.\
**Gender.** Similar to AltX and EgoX in prior models, both an alter's
gender (b = -0.0903, SE = 0.3372) and an ego's gender (b = 0.2993, SE =
0.4934) still have insignificant effects. The parameter SameX shows a
significant positive estimate (b = 0.6820, SE = 0.2804) which barely
differs from the two prior models. This suggests that adding the
covariate 'H-Index' did not prompt any major changes in scholars'
preferences for same-gendered collaborators.\
**Department.** While an alter's departmen**t** (AltX) remained
insignificant to the collaboration patterns, the estimates for an ego's
department changed incomparison to the prior model. In this model, egoX
seems to be of significant relevance (b = -1.4652, SE = 0.6963), with
its negative estimate indicating that, once effects of h-index are
controlled for, scholars from the political science (coded as 1)
department have a lower tendency to initiate collaborations than
scholars from the sociology department (coded as 0).\
Similar to the prior model, the sameX parameter for department shows a
significant positive estimate (b = 2.6327, SE = 0.6080) scholars are
still more likely to collaborate with others from the same department.\
**H-Index**. Both an alter's and an ego's H-Index seem to be relevant in
collaboration preferences. The significant and positive estimateof AltX
(b = 0.0829, SE = 0.0364) suggest that scholars with higher h-index are
more preferred as collaborators. The egoX estimate is significant and
negative (b = -0.1591, SE = 0.0744) suggesting that scholars with higher
h-index are less likely to initiate collaborations.\
The results partially support **H4**, as the h-index is shown to
influence collaboration patterns, with prestigious scholars being more
likely to attract collaborations as alters but less likely to initiate
them as egos. However, gender homophily remains significant even after
controlling for the h-index, suggesting that prestige does not
substantially weaken the influence of gender homophily in co-authorship
networks.

### 2024-11-25 finalintro.Rmd
firstfinalpushmaybe

---
title: "Introduction"
author: "Verooo"
date: "2024-11-23"
output: html_document
bibliography: references.bib
csl: apa.csl
---
```{=html}
<style>
body {
text-align: left;
line-height: 1.5;
}

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

>>>>

---
title: "Introduction"
author: "Verooo"
date: "2024-11-23"
output: html_document
bibliography: references.bib
csl: apa.csl
---

```{=html}
<style>
body {
text-align: left;
line-height: 1.5;
}

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

------------------------------------------------------------------------

# Introduction

## Contextualization

Collaboration has become a cornerstone of modern academic practice, with scholars increasingly working together to leverage diverse expertise, share resources, and enhance their research output. Co-authorship, in particular, is a key indicator of collaboration, often reflecting the dynamics of professional relationships and the distribution of academic opportunities. As Cheong (@Cheong2009) highlights, collaboration enables researchers to combine intellectual and institutional resources, contributing significantly to their scholarly success. However, collaboration is not free from social patterns and biases. Academic networks often mirror broader societal inequalities, with persistent disparities in opportunities and recognition. One such disparity id gender inequality, which continues to shape interactions, opportunities, and outcomes within academia.

## Research Problem and Goal

In this study, I examine the role of gender homophily in shaping co-authorship networks in the Sociology and Political Science Departments at Radboud University. Gender homophily refers to the tendency of individuals to collaborate with others of the same gender. This phenomenon can influence who gains access to collaboration opportunities, thereby shaping academic trajectories and reinforcing existing inequalities. Understanding these dynamics is essential to uncovering how collaboration networks may perpetuate or lessen gender-based disparities in academia.

## Prior findings

### Gender and Collaboration

Previous research has consistently shown that gender plays a significant role in academic collaboration. Horta and colleagues (@Horta2022) reveal that male researchers are more likely to collaborate with other men, while female researchers exhibit a weaker but still present tendency towards homophily. Kwiek and Roszka (@Kwiek2021) found that such patterns contributes to the persistence of gender imbalances in co-authorship, where all-male teams remain more common than all-female teams. These trends are particularly concerning in fields such as Sociology and Political Science, where collaboration is vital for career advancement (@Cheong2009). The structure of these networks affects not only who collaborates but also how academic recognition and resources are distributed (@Cheong2009). For example, male-dominated networks may provide better access to high-impact collaborations and grant opportunities for men, further perpetuating the gender gap in academic successes.

### Prestige and Collaboration

Another crucial factor influencing collaboration id academic prestige. Scholars with higher prestige, indicated by metrics such as the h-index or publication in top-tier journals, are generally more sought after as collaborators. Existing research suggests that prestige might interact with gender homophily in significant ways. Bellotti et al. (@Bellotti2022) found that prestigious male scholars often form networks that exclude women, while McDowell and colleagues (@Mcdowell2007) also observed that female scholars, despite demonstrating similar academic competence, face greater challenges in accessing high-prestige collaborative networks. In this context, it is crucial to explore whether the tendency for gender-based collaboration shifts when prestige factors come into play. Specifically, do scholars' h-indeces exacerbate gender homophily in academic networks?

## Research Focus and Questions

This study aims to analyze gender homophily within the co-authorship networks of the Sociology and Political Science Departments at Radboud University, focusing on the role of prestige. The research is guided by the following questions:

-   ***RQ1a:** To what extent does gender homophily occur across all co-publications of the staff in the Sociology and Political Science Departments at Radboud University?*

-   ***RQ1b:** To what extent do the two departments differ in their levels of gender homophily?*

These questions focus on descriptive patterns, providing a macro-level view of gender clustering in academic networks.

-   ***RQ2:** To what extent can gender homophily in collaboration networks be explained by preferences at the dyadic level, controlling for structural network and individual-level covariates?*

This question shifts the focus to the micro-level, examining whether individual choices and structural factors drive homophily in co-authorship.

-   **RQ3:** How does the prestige of scholars, measured by scholars' h-index, influence gender homophily in co-authorship networks?

This question explores whether prestige interacts with gender homophily, shedding light on how academic recognition impacts collaborative behaviors. Specifically, it investigates whether male and female scholars with similar prestige levels experience different patterns of collaboration and whether prestige reduces or amplifies gender clustering.

## Relevance and Contribution

This study contributes to the growing body of literature on gender inequalities in academia by focusing on the intersection of gender and academic prestige in shaping collaboration networks. By applying a social network perspective, it offers insights into how gender-based preferences influence co-authorship patterns and how these dynamics vary between disciplines. Moreover, by exploring the role of journal prestige, this research highlights how structural factors may intersect with social biases to shapw academic opportunities, providing a nuanced understanding of the mechanisms underlying gender disparities in academia.

------------------------------------------------------------------------

# References
